On Devon Allen's Disqualification at the 2022 Track and Field World Championship

### 2023-04-14 Manuscript/manuscript.tex
comments/suggestons

  \jy{Avoid references in abstract.}

  \jy{This is only the background. An abstract also needs to summarize the contributions.}
%\bigskip
\noindent\sc{Keywords}:
Reaction Time, Seiko,

@@

\jy{Not a good opening. Could start by telling the story, recapping the major
  events and discussions on the internet. Then dig into the relevant
  literature. Finally outline the contributions of this work.
}

@@

%1
After coming in third in the 110 meter hurdles at the United States Track and 
Field Championships in Eugene Oregon at the end of June, Allen was expected to 
be a contender for the World Championships, which were scheduled for two months 
later, also in Eugene. At the World Championships, Allen made it through the 
heats and semifinals to reach the finals while posting reaction times of 0.123
in the heats and 0.101 in the semifinal.  Allen, who attended the University
of Oregon (Location of US World Championship and the World Championship) was
considered to be a favorite for the 110 meter hurdle after running 12.84 seconds
in early June; only 0.04 seconds away from the world record \citep{Preview}.
Thus his disqualification by reacting in 0.099 seconds instead of 0.01 was met
with shock from both Allen and fans alike.  NBC, who was broadcasting the
championships in the United States, later uploaded a video to their YouTube
channel that show the disqualification and can be found here:
\url{https://www.youtube.com/watch?v=D6NXTMo-1yM}. The crowd erupted with jeers when
they found out that Allen was disqualified, angry at the result.  Had Allen 
reacted just 0.001 seconds slower he would have been allowed to compete and been 
spared waiting another two years until the next World Track and Field Championship.
For track and field athletes, they usually only get the chance to prove 
themselves to sponsors every two years, meaning that races such as the finals
of the World Championship is an impactful way to earn a living. An 
investigation into reaction times is also important because reaction times have
been shown to affect the overall time of the race \citep{delalija2008reaction}.
Thus, faster reaction times lead to faster overall times; which is the goal of
every sprinter who steps onto the track.

%2
Outside the blog posts from LetsRun.com, there have not been any published
papers regarding Devon Allen's disqualification in particular.  This is not
particularly surprising given that his disqualification occurred only last August.
Letsrun is a website that is part message board and part news.  The message board,
which functions similarly to Reddit is very active during important running
events, which include during the finals of major events such as
the Olympics and the World Track and Field Championships.  However, it is the
news section of the website that generated articles such as ``Was Devon Allen
Screwed? There's At Least a 99\% That He Was" and ``The Data Keeps Pouring In and
It Continues to Look Bad For World Athletics and Great For Devon Allen".  Robert
Johnson, who created and runs the Letsrun website argued that there was an error
with the timing equipment at the 2022 Track and Field World Championship and
cited graphs, statistics such as median reaction time, and compared the reaction
times at the 2022 Track and Field World Championship.

%3
In 2009, the International Association of Athletics Federations (Also referred to
as IAAF or World Athletics), which is the governing board for international running
competitions, commissioned a study into changing the reaction time barrier.
The paper analyzed sprinters reaction times and the biological processes 
that went into reacting to the starting gun and concluded 
that athletes can react in as low as 0.08 seconds.  One study by Pain looked at
reaction times for nine male athletes and analyzed their reaction times.  The
fastest athlete of the nine, had on average (even with their two fastest time 
removed) a reaction time of 0.087 seconds and a standard deviation of 0.004
seconds.  Two other athletes were able to react similarly when under certain
muscular tightening conditions \citep{pain2007sprint}. Another study commissioned
by IAAF concluded that the disqualification barrier be decreased to allow sprinters to
react faster, and suggested 0.08 or 0.085 seconds as possible new thresholds.
Additionally, it was recommended to IAAF that high speed cameras be used to
change the reaction time criteria from pushing off on the block to body movement.
\citep{komi2009iaaf} This could provide a more accurate representation of 
reaction time, and thus the possibility of a false disqualification would be lowered.
The possibility of using a video-based medium has been proved to be a reliable
measure of reaction time \citep{mudric2015evaluation}.
A 2012 follow-up study to Komi's by Pilianidis found
significant differences between reaction times at several world championships
between 1997 and 2009 for the men's 110 meter hurdles.  Another important
conclusion was that for reaction times did not decrease over the twelve years
covered by the research in the study \citep{pilianidis2012start}.  This contradicts
what was found in a 2021 study that concluded the year of competition has a
significant effect on reaction times \citep{zhang2021correlation}.  Zhang's
analysis looked at data from the 2011-2019 World Championships across multiple
sprint events and concluded that from 2011 to 2019, athletes reaction times got
faster. However, not  all studies have been in favor of a decrease in the
reaction time barrier. Brosnan in 2016 argues for increasing the reaction time 
barrier to 0.115 seconds, and cited evidence from European Championships and 
World Championships in his article \citep{brosnan2017effects}.  A study looking 
at the 2008 Olympics likewise supported the conclusion that athletes could not
react in as little as 0.1 seconds \citep{lipps2011implications}.  Another study
concluded that visual reaction times to the starting gun may be faster by 0.007
seconds than the IAAF method, which is from force on block starts 
\citep{holmes2018method}.  While this may seem trivial, Allen was disqualified
only by 0.01 seconds. It is also worth noting that one study looking at data 
from various IAAF meets found that the fastest reaction times for men were found
between the ages of 26 and 29 \citep{tonnessen2013reaction}.  This is notable 
because Allen was 27 years old at the 2022 Track and Field World Championship 
and in his physical peak.

@@

%4
One of the first objectives of the paper is to try to model the data, which is
the reaction times of every non-disqualification since the 1999 World Championships. 
The model we use is a linear model, following a gamma distribution with a random
effect, which will be referred to as the year effect.  The year effect is
reflective of how significant the changes in the reaction time is for each venue,
i.e. the venue effect is  responsible for the variation between years.  
Another objective of this paper is to compare data for athletes who have competed
at additional competitions to see how their reaction times compare from previous
competitions to this one. First we can look at the reaction times of athletes who
competed at the World Track and Field Championship in both 2019 and 2022 and
compare the times, and secondly we can look at how athletes performed compared
to how they did at their national/qualifiers meet. In order to reduce any chances
of athletes improving reaction time, we can also look at data from qualifying meets.
Almost every athlete had to qualify by proving they are elite at the national level, 
usually by coming in at least fourth. Thus, data was compiled on athletes who competed
at the 2022 World Track and Field Championships on what these athletes also ran 
at their country's national championship. If there was any argument that athletes
have improved their reaction time from 2019 to 2022, it should now be considered
nearly null, as the national competitions took place between June and July 2022 
and the World Championships were in August. The equipment used to measure reaction
time cannot be guaranteed to be the same from national to international
competition, but the time span between meets has been reduced considerably.

@@

\subsection{GLMM}
Data was taken from the World Athletics and covers the men's 110 meter hurdles 
and the women's 100 meter hurdles from 1999 to 2022.  The variable of prediction 
interest is reaction time, and the predicting variables are year, stage of 
competition (heats, semifinal, final), total time, and gender.  
The data can be found in the appendix~\ref{sec:Appendix} 
Factors such as name, country, and identification number were
not considered important for the purposes of this research besides the times for
Devon Allen and the other United States athletes mentioned above.  The variable "Stage"
refers to whether the observation occurred during the "heats" (Preliminary round) which
is denoted by "H" in the data, "S" for "semifinals" (Of which there are either 2 or 3 
heats each year), and "F" for finals. The variable "Gender" has levels "M" and "F" 
which stand for "male" and "female" respectively.  The variable "Batch" takes a 
numerical value and every heat of every round from 1999 to 2022 was assigned it's
own number.  This was added because of suspected issues that reaction time may be
correlated with other runners reacting quickly (one runner reacting quickly may
result in other runners reacting quickly).  There is no statistical significance to
how the batches were labeled, they were ordered chronologically by meet but
reverse chronologically by year (Batch 1 is a preliminary Heat of 2022 and Batch
111 corresponds to the 1999 finals). Included are
box plots of the reaction times of the men's finals and the men's semifinals and
finals pooled together.  We will refer to the latter as the "pooled data"
throughout the paper.  The heats or preliminary rounds were not studied because
athletes react faster as the competition progresses.  Athletes react faster in
the finals than the semifinals which are faster than the heats 
\citep{zhang2021correlation}.  If we want to determine if Allen's reaction time
is reasonable, then we thus need to look at the men's final data, with the men's
pooled data being included in order to increase the sample size.  In 2022, there
were only five data points due to two disqualifications and one athlete who 
qualified but did not compete.
The data that looked at in the paper is best visualized using a box plot as show 
in Figure~\ref{fig:Boxplots}.  It is clear that the 2022 box plot for both the
men's finals and semifinals data is lower (indicating a faster reaction time)
than in previous years.  The median for the men's pooled data in 2022 was 0.129
and the median for the men's final data in 2022 was 0.126.  For comparison,
Brosnan found reaction times of 0.156 seconds for men when looking at data from
1999 to 2014 \citep{brosnan2017effects}.
%Fix quotations
%Dont need the titles will add captions, use width =5, height = 3 in aspect ratio

@@

\subsection{Rank based Comparisons}
This section was motivated from a qualitative analysis that started by looking
at how United States athletes reacted at the 2022 United States Track and Field
Championships. From June 23-26 2022, USA held its Track
and Field Championship to decide who to send to the World Championships at Hayward 
Field, the exact same venue as used for the World Track and Field Championship a month 
later.  Thus, a baseline is able to be established for the four USA athletes who 
competed in at least one round of both events: Trey Cunningham, Daniel Roberts, 
Grant Holloway, and Devon Allen. Every athlete reacted faster in all the World 
Track and Field Championship races compared to the USA Track and Field 
Championship races. Devon Allen raced three times at the USA Track and
Field Championship and three times at the World Track and Field  Championship. 
His reaction times were significantly faster in the World Track and
Field Championship compared to the USA Track and Field Championship as were his
teammates.  This comparison suggests that the timing system used between the two
meets may have been meaningfully different to the point where Devon Allen was
disqualified more because of faulty equipment rather than because he reacted
too quickly.  This table below highlights the difference in reaction times for
Devon Allen, Trey Cunningham, Grant Holloway, and Daniel Roberts. "USA" denotes
the USA Track and Field Championships, "World" denotes the World Track and Field
Championships, "N/A" was used for any athlete that did not compete and thus
did not register a reaction time.  All numbers listed in the table are the reaction
time for the respective athletes in seconds.

\begin{center}
  \begin{tabular}{|c c c c c c c|} 
   \toprule
   Athlete & USA H & USA S & USA F & World H & World S & World F \\ [0.5ex] 
   \midrule
   Devon Allen & 0.201 & 0.153 & 0.160 & 0.123 & 0.101 & 0.099 \\ 
   Trey Cunningham & 0.186 & 0.185 & 0.182 & 0.115 & 0.120 & 0.109 \\
   Grant Holloway & 0.192 & 0.190 & N/A & 0.147 & 0.128 & 0.124 \\
   Daniel Roberts & 0.181 & 0.200 & 0.183 & 0.179 & N/A & N/A \\ [0.5ex]
   \bottomrule
  \end{tabular}
  \end{center}

@@

\subsection{Rank-based comparison for Clustered}

@@

The test statistic of the DS method of the Wilcoxon rank-sum statistic is:
$Z = \frac{S - E(S)}{\sqrt{\hat{\text{Var}}(S)}}$ where 
$W^* = \frac{1}{N} + \sum_{i=1}^{N} \delta_{i}^{*} R_{i}^{*}$, $R_{i}^{*}$ is
the rank of $X_{i}^{*}$.
The asymptomatic distribution of the Wilcoxon rank-sum statistic was found by
\citet{datta2005rank} and the R package "clusrank" allows for its implementation.
\citep{jiang2017wilcoxon}.

Using the "clusWilcox.test" function, a grouped and clustered Wilcoxon rank sum
test was performed with the cluster set to be the Athlete and group set to Year.
The method was set to be "DS" which is used for the Data and Satten method, which
is preferable when the cluster sizes are informative \citep{jiang2017wilcoxon}.  We do not care that some
athletes made it further in either competition than another or that it should
not matter how elite an athlete is at the 110 meter hurdles once we get to the
national international competition level.  Cluster sizes vary from 2 observations
to 6.

\section{Results} \label{sec:Results}
\jy{summarize your findings; discuss limitations and possible future works.}

\subsection{GLMM Results}
%Section 4.1: Gamma model results show the fitting with and without 2022,
%talk about p values and thresholds from the simulations (do not include the
%results in the methods section). 
At a size of n=10,000,000, the calculated probability of a reaction
time being below 0.1 seconds was as follows for the various different data sets:

\begin{center}
  \begin{tabular}{|c c|} 
   \hline
   Data Set & Probability RxnTime < 0.1 \\ 
   \hline\hline
   all finals gamma log & 0.000856 \\ 
   \hline
   old finals gamma log & 0.000439 \\
   \hline
   all pooled gamma log & 0.006865 \\
   \hline
   old pooled gamma log & 0.005761 \\
  \end{tabular}
  \end{center}

@@

%Citation needed in this section
Since 1985 Seiko Holding Corporations has served every World Athletics Championship
as the official timer \citep{Seiko}.  Since 1985, the technology and the ability
to accurately predict measurements: long jump distances, false starts, total time,
reaction time, etc. have all dramatically improved.  Seiko did not start tracking
reaction time as an official measurement until 1999 in the men's and women's 100 meter
hurdles.  Seiko regularly updates their equipment so that they provide cutting edge
technology to the World Track and Field Championships, the highest stakes in the world
of running outside of the Olympics. Seiko's technology for detecting reaction 
times relies on the pressure that athletes
exert on the plate when they push off.  Their systems measure the time differential
between when the "gun" goes off to start the race and when the pressure changes 
\citep{Seiko}  If an athlete has a reaction time under 0.10 seconds, it is deemed a 
false start as it is considered that no athlete can react so quickly 
\citep{Seiko-Timing}.  Thus, a reaction time of for example 0.05, suggests that 
the athlete predicted the gun and it was luck that caused their abnormally high 
time.  At the World Championship level, World Athletics and Seiko want to remove 
that element of luck and thus impose the 0.1 second barrier. In 2013, Seiko 
upgraded their timing equipment for sprinting events (100 meter hurdles falls 
under this category) \citep{WorldAthletics_2013}.  Since 1999, the highest reaction 
times in the men's 110-meter hurdle were recorded in 2013.  That is not to say that the higher 
times were caused by Seiko's equipment, but the two may be related.  It is worth noting that prior
to 2022, Seiko again upgraded its technology but for its jump management system.

\subsection{Rank Comparison Results}
%Section 4.2 is motivated by seiko timing and look at same athletes across
%multiple types of competition: 2019 and national

@@

When the clusrank rank sum test was performed, the "ds" method produced a z score of
2.9751 and p value of 0.001464.  This value is for a one-sided test and show that
the mean reaction time on average in 2019 was larger than the mean reaction time
in 2022 for athletes who competed in both championships at an alpha level of 0.01.
This provides good evidence that athletes who competed in both 2019 and 2022 got
faster during that timespan.

Repeating the data collection steps and
again performing a clustered analysis for a Wilcoxon ranked sum test
results in two new z scores and p values.  The "ds" method produced a z score of
3.6069 and p value of 0.000155.  This result is highly significant and shows that
the times at the national stage of competition were significantly higher than
the reaction times at the World Championships even for an alpha level of 0.001.

>>>>

\bigskip
\noindent{\sc Keywords}:
Reaction Time, Seiko,

@@

@@

\jy{could condense a bit}
%1
After coming in third in the 110 meter hurdles at the United States Track and 
Field Championships in Eugene Oregon at the end of June, Allen was expected to 
be a contender for the World Championships, which were scheduled for two months 
later, also in Eugene. At the World Championships, Allen made it through the 
heats and semifinals to reach the finals while posting reaction times of 0.123
in the heats and 0.101 in the semifinal.  Allen, who attended the University
of Oregon (Location of US World Championship and the World Championship) was
considered to be a favorite for the 110 meter hurdle after running 12.84 seconds
in early June; only 0.04 seconds away from the world record \citep{Preview}.
Thus his disqualification by reacting in 0.099 seconds instead of 0.01 was met
with shock from both Allen and fans alike.  NBC, who was broadcasting the
championships in the United States, later uploaded a video to their YouTube
channel that show the disqualification and can be found here:
\url{https://www.youtube.com/watch?v=D6NXTMo-1yM}. The crowd erupted with jeers when
they found out that Allen was disqualified, angry at the result.  Had Allen 
reacted just 0.001 seconds slower he would have been allowed to compete and been 
spared waiting another two years until the next World Track and Field Championship.
For track and field athletes, they usually only get the chance to prove 
themselves to sponsors every two years, meaning that races such as the finals
of the World Championship is an impactful way to earn a living. An 
investigation into reaction times is also important because reaction times have
been shown to affect the overall time of the race \citep{delalija2008reaction}.
Thus, faster reaction times lead to faster overall times; which is the goal of
every sprinter who steps onto the track.

%2
Outside the blog posts from \url{www.LetsRun.com}, there have not been any published
papers regarding Devon Allen's disqualification in particular.  This is not
particularly surprising given that his disqualification occurred only last August.
Letsrun is a website that is part message board and part news.  The message board,
which functions similarly to Reddit is very active during important running
events, which include during the finals of major events such as
the Olympics and the World Track and Field Championships.  However, it is the
news section of the website that generated articles such as ``Was Devon Allen
Screwed? There's At Least a 99\% That He Was" and ``The Data Keeps Pouring In and
It Continues to Look Bad For World Athletics and Great For Devon Allen".  Robert
Johnson, who created and runs the Letsrun website argued that there was an error
with the timing equipment at the 2022 Track and Field World Championship and
cited graphs, statistics such as median reaction time, and compared the reaction
times at the 2022 Track and Field World Championship.

\jy{break into two paragraphs}
%3
In 2009, the International Association of Athletics Federations (Also referred to
as IAAF or World Athletics), which is the governing board for international running
competitions, commissioned a study into changing the reaction time barrier \citep{reference?}.
The paper analyzed sprinters reaction times and the biological processes 
that went into reacting to the starting gun and concluded 
that athletes can react in as low as 0.08 seconds.   \citet{pain2007sprint} looked at
reaction times for nine male athletes and analyzed their reaction times.  The
fastest athlete of the nine, had on average (even with their two fastest time 
removed) a reaction time of 0.087 seconds and a standard deviation of 0.004
seconds.  Two other athletes were able to react similarly when under certain
muscular tightening conditions. Another study commissioned
\jy{Is this the same study at the beginning of this paragraph?}
by IAAF concluded that the disqualification barrier be decreased to allow sprinters to
react faster, and suggested 0.08 or 0.085 seconds as possible new thresholds.
Additionally, it was recommended to IAAF that high speed cameras be used to
change the reaction time criteria from pushing off on the block to body movement.
\citep{komi2009iaaf}. This could provide a more accurate representation of 
reaction time, and thus the possibility of a false disqualification would be lowered.
The possibility of using a video-based medium has been proved to be a reliable
measure of reaction time \citep{mudric2015evaluation}.
A 2012 follow-up study to Komi's by Pilianidis found
significant differences between reaction times at several world championships
between 1997 and 2009 for the men's 110 meter hurdles.  Another important
conclusion was that for reaction times did not decrease over the twelve years
covered by the research in the study \citep{pilianidis2012start}.  This contradicts
what was found in a 2021 study that concluded the year of competition has a
significant effect on reaction times \citep{zhang2021correlation}.  Zhang's
\jy{Not Zhang's. There are coauthors. Consider using citet when appropriate in this paragraph.}
analysis looked at data from the 2011-2019 World Championships across multiple
sprint events and concluded that from 2011 to 2019, athletes reaction times got
faster. However, not  all studies have been in favor of a decrease in the
reaction time barrier. Brosnan in 2016 argues for increasing the reaction time 
barrier to 0.115 seconds, and cited evidence from European Championships and 
World Championships in his article \citep{brosnan2017effects}.  A study looking 
at the 2008 Olympics likewise supported the conclusion that athletes could not
react in as little as 0.1 seconds \citep{lipps2011implications}.  Another study
concluded that visual reaction times to the starting gun may be faster by 0.007
seconds than the IAAF method, which is from force on block starts 
\citep{holmes2018method}.  While this may seem trivial, Allen was disqualified
only by 0.01 seconds. It is also worth noting that one study looking at data 
from various IAAF meets found that the fastest reaction times for men were found
between the ages of 26 and 29 \citep{tonnessen2013reaction}.  This is notable 
because Allen was 27 years old at the 2022 Track and Field World Championship 
and in his physical peak.

@@

%4
\jy{I think you have one objective but two methods to approach it.}
One of the first objectives of the paper is to try to model the data, which is
the reaction times of every non-disqualification since the 1999 World Championships. 
The model we use is a linear model, following a gamma distribution with a random
effect, which will be referred to as the year effect.  The year effect is
reflective of how significant the changes in the reaction time is for each venue,
i.e. the venue effect is  responsible for the variation between years.  
Another objective of this paper is to compare data for athletes who have competed
at additional competitions to see how their reaction times compare from previous
competitions to this one. First we can look at the reaction times of athletes who
competed at the World Track and Field Championship in both 2019 and 2022 and
compare the times, and secondly we can look at how athletes performed compared
to how they did at their national/qualifiers meet. In order to reduce any chances
of athletes improving reaction time, we can also look at data from qualifying meets.
Almost every athlete had to qualify by proving they are elite at the national level, 
usually by coming in at least fourth. Thus, data was compiled on athletes who competed
at the 2022 World Track and Field Championships on what these athletes also ran 
at their country's national championship. If there was any argument that athletes
have improved their reaction time from 2019 to 2022, it should now be considered
nearly null, as the national competitions took place between June and July 2022 
and the World Championships were in August. The equipment used to measure reaction
time cannot be guaranteed to be the same from national to international
competition, but the time span between meets has been reduced considerably.

@@

\subsection{World Championship}
Data was taken from the World Athletics and covers the men's 110 meter hurdles 
and the women's 100 meter hurdles from 1999 to 2022.  The variable of prediction 
interest is reaction time, and the predicting variables are year, stage of 
competition (heats, semifinal, final), total time, and gender.  
The data can be found in the appendix~\ref{sec:Appendix} 
Factors such as name, country, and identification number were
not considered important for the purposes of this research besides the times for
Devon Allen and the other United States athletes mentioned above.  The variable ``Stage"
refers to whether the observation occurred during the ``heats" (Preliminary round) which
is denoted by ``H" in the data, ``S" for ``semifinals" (Of which there are either 2 or 3 
heats each year), and ``F" for finals. The variable ``Gender" has levels ``M" and ``F" 
which stand for ``male" and ``female" respectively.  The variable ``Batch" takes a 
numerical value and every heat of every round from 1999 to 2022 was assigned it's
own number.  This was added because of suspected issues that reaction time may be
correlated with other runners reacting quickly (one runner reacting quickly may
result in other runners reacting quickly).  There is no statistical significance to
how the batches were labeled, they were ordered chronologically by meet but
reverse chronologically by year (Batch 1 is a preliminary Heat of 2022 and Batch
111 corresponds to the 1999 finals). Included are
box plots of the reaction times of the men's finals and the men's semifinals and
finals pooled together.  We will refer to the latter as the ``pooled data"
throughout the paper.  The heats or preliminary rounds were not studied because
athletes react faster as the competition progresses.  Athletes react faster in
the finals than the semifinals which are faster than the heats 
\citep{zhang2021correlation}.  If we want to determine if Allen's reaction time
is reasonable, then we thus need to look at the men's final data, with the men's
pooled data being included in order to increase the sample size.  In 2022, there
were only five data points due to two disqualifications and one athlete who 
qualified but did not compete.
The data that looked at in the paper is best visualized using a box plot as show 
in Figure~\ref{fig:Boxplots}.  It is clear that the 2022 box plot for both the
men's finals and semifinals data is lower (indicating a faster reaction time)
than in previous years.  The median for the men's pooled data in 2022 was 0.129
and the median for the men's final data in 2022 was 0.126.  For comparison,
Brosnan found reaction times of 0.156 seconds for men when looking at data from
1999 to 2014 \citep{brosnan2017effects}.
%Fix quotations
%Dont need the titles will add captions, use width =5, height = 3 in aspect ratio

@@

\subsection{Beyond World Championship}
This section was motivated from a qualitative analysis that started by looking
at how United States athletes reacted at the 2022 United States Track and Field
Championships. From June 23-26 2022, USA held its Track
and Field Championship to decide who to send to the World Championships at Hayward 
Field, the exact same venue as used for the World Track and Field Championship a month 
later.  Thus, a baseline is able to be established for the four USA athletes who 
competed in at least one round of both events: Trey Cunningham, Daniel Roberts, 
Grant Holloway, and Devon Allen. Every athlete reacted faster in all the World 
Track and Field Championship races compared to the USA Track and Field 
Championship races. Devon Allen raced three times at the USA Track and
Field Championship and three times at the World Track and Field  Championship. 
His reaction times were significantly faster in the World Track and
Field Championship compared to the USA Track and Field Championship as were his
teammates.  This comparison suggests that the timing system used between the two
meets may have been meaningfully different to the point where Devon Allen was
disqualified more because of faulty equipment rather than because he reacted
\jy{cross-reference to the table by its label}
too quickly.  This table below highlights the difference in reaction times for
Devon Allen, Trey Cunningham, Grant Holloway, and Daniel Roberts. ``USA" denotes
the USA Track and Field Championships, ``World" denotes the World Track and Field
Championships, ``N/A" was used for any athlete that did not compete and thus
did not register a reaction time.  All numbers listed in the table are the reaction
time for the respective athletes in seconds.

\begin{center}
  \begin{tabular}{c c c c c c c} 
   \toprule
   Athlete & USA H & USA S & USA F & World H & World S & World F \\ [0.5ex] 
   \midrule
   Devon Allen & 0.201 & 0.153 & 0.160 & 0.123 & 0.101 & 0.099 \\ 
   Trey Cunningham & 0.186 & 0.185 & 0.182 & 0.115 & 0.120 & 0.109 \\
   Grant Holloway & 0.192 & 0.190 & N/A & 0.147 & 0.128 & 0.124 \\
   Daniel Roberts & 0.181 & 0.200 & 0.183 & 0.179 & N/A & N/A \\ [0.5ex]
   \bottomrule
  \end{tabular}
  \end{center}

@@

\subsection{Rank-based Comparison}

@@

\jy{Undefined notations}
The test statistic of the DS method of the Wilcoxon rank-sum statistic is:
$Z = \frac{S - E(S)}{\sqrt{\hat{\text{Var}}(S)}}$ where 
$W^* = \frac{1}{N} + \sum_{i=1}^{N} \delta_{i}^{*} R_{i}^{*}$, $R_{i}^{*}$ is
the rank of $X_{i}^{*}$.
The asymptomatic distribution of the Wilcoxon rank-sum statistic was found by
\citet{datta2005rank} and the R package ``clusrank" allows for its implementation.
\citep{jiang2017wilcoxon}.

Using the \texttt{clusWilcox.test()} function, a Wilcoxon rank sum
test was performed with the cluster set to be the Athlete and group set to Year
using the DS method. The DS method is preferable when the cluster sizes are
informative \citep{jiang2017wilcoxon}. \jy{explain. More data means faster
  reaction time?}
We do not care that some
athletes made it further in either competition than another or that it should
not matter how elite an athlete is at the 110 meter hurdles once we get to the
national international competition level.  Cluster sizes vary from 2 observations
to 6.

\section{Results} \label{sec:Results}

\subsection{GLMM}
%Section 4.1: Gamma model results show the fitting with and without 2022,
%talk about p values and thresholds from the simulations (do not include the
%results in the methods section).

\jy{present the estimated parameters first before deriving the p-value using the
  estimated parameters. The difference in variance of the random effects using
  different data motivated the rank-based study.}

At a size of $n=10,000,000$, the calculated probability of a reaction
time being below 0.1 seconds was as follows for the various different data sets:
\jy{label the table and reference to it; clean it up}
\begin{center}
  \begin{tabular}{|c c|} 
   \hline
   Data Set & Probability RxnTime < 0.1 \\ 
   \hline\hline
   all finals gamma log & 0.000856 \\ 
   \hline
   old finals gamma log & 0.000439 \\
   \hline
   all pooled gamma log & 0.006865 \\
   \hline
   old pooled gamma log & 0.005761 \\
  \end{tabular}
  \end{center}

@@

%Citation needed in this section
Since 1985 Seiko Holding Corporations has served every World Athletics Championship
as the official timer \citep{Seiko}.  Since 1985, the technology and the ability
to accurately predict measurements: long jump distances, false starts, total time,
reaction time, etc. have all dramatically improved.  Seiko did not start tracking
reaction time as an official measurement until 1999 in the men's and women's 100 meter
hurdles.  Seiko regularly updates their equipment so that they provide cutting edge
technology to the World Track and Field Championships, the highest stakes in the world
of running outside of the Olympics. Seiko's technology for detecting reaction 
times relies on the pressure that athletes
exert on the plate when they push off.  Their systems measure the time differential
between when the ``gun" goes off to start the race and when the pressure changes 
\citep{Seiko}  If an athlete has a reaction time under 0.10 seconds, it is deemed a 
false start as it is considered that no athlete can react so quickly 
\citep{Seiko-Timing}.  Thus, a reaction time of for example 0.05, suggests that 
the athlete predicted the gun and it was luck that caused their abnormally high 
time.  At the World Championship level, World Athletics and Seiko want to remove 
that element of luck and thus impose the 0.1 second barrier. In 2013, Seiko 
upgraded their timing equipment for sprinting events (100 meter hurdles falls 
under this category) \citep{WorldAthletics_2013}.  Since 1999, the highest reaction 
times in the men's 110-meter hurdle were recorded in 2013.  That is not to say that the higher 
times were caused by Seiko's equipment, but the two may be related.  It is worth noting that prior
to 2022, Seiko again upgraded its technology but for its jump management system.

\subsection{Rank-based Comparison}
%Section 4.2 is motivated by seiko timing and look at same athletes across
%multiple types of competition: 2019 and national

@@

When the clusrank rank sum test was performed, the ``ds" method produced a z score of
2.9751 and p value of 0.001464.  This value is for a one-sided test and show that
the mean reaction time on average in 2019 was larger than the mean reaction time
in 2022 for athletes who competed in both championships at an alpha level of 0.01.
This provides good evidence that athletes who competed in both 2019 and 2022 got
faster during that timespan.

Repeating the data collection steps and
again performing a clustered analysis for a Wilcoxon ranked sum test
results in two new z scores and p values.  The ``ds" method produced a z score of
3.6069 and p value of 0.000155.  This result is highly significant and shows that
the times at the national stage of competition were significantly higher than
the reaction times at the World Championships even for an alpha level of 0.001.
